|

### 2022-07-02 article/article.Rmd
small edits

---
title:  |
        | Legislature Size and Welfare: Evidence from Brazil
author: 
  - Umberto Mignozzetti[^umberto]
  - Gabriel Cepaluni[^gabriel]
  - Danilo Freire[^danilo]
date: May 28, 2022
thanks: We would like to thank Victor Araujo, Eric Arias, Mark Bennister, Taylor Boas, Natalia Bueno, Nick Cowen, Andrea Freitas, Saad Gulzar, Shoaib Jillani, James Mahon, Lucas Mingardi, Lucas Novaes, Ingrid Oliveira, Catarina Roman, Nelson Ruiz, David Skarbek, Mike Slaven, Denis Stukal, the AJPS Editorial Team, and three anonymous reviewers for their valuable comments. We also thank the participants of seminars and conferences at the CEPESP-FGV, CUNY, INSPER, IPSA, IR-FGV, LACEA, LASA, Lincoln, MPSA, Notre Dame, NYU, PUC Chile, SPSA, and Uppsala. All remaining errors are the sole responsibility of the authors. The authors thank FAPESP for their financial support, grant number 2017/07290-5. Replication materials are available at <https://github.com/umbertomig/legislature-size-welfare>.
abstract: "How does legislature size impact public service provision? Despite the importance of institutional design for democratic governance, the effect of legislative features on citizen welfare remains little understood. In this paper, we use a formal model to show that increasing legislature size improves public goods delivery. We argue that changes in bargaining costs depend on whether additional legislators share the executive's party affiliation: More opposition members reduce the equilibrium public goods provision, while more government-aligned members increase it. We test this theory by exploiting sharp discontinuities in city council size in Brazil. We show that an additional city councilor has a 91% chance of belonging to the mayoral coalition, and this significantly improves primary school enrollment and infant mortality rates. To explore possible mechanisms, we surveyed 174 former city councilors and analyzed 346,553 bills proposed between 2005 and 2008. This paper has implications for the design of representative institutions."
abstractspacing: double
keywords: legislature size; legislative bargaining; local public goods provision; regression discontinuity design; Brazil
jelcodes: H41; H80; I31; I38
fontsize: 12pt
margin: 2cm
urlcolor: darkblue
linkcolor: Mahogany
citecolor: Mahogany
spacing: double
papersize: a4paper
bibliography: references.bib
biblio-style: apalike
output:
  pdf_document:
    citation_package: natbib
    fig_caption: yes
    number_sections: yes
    keep_tex: no
    toc: no
    toc_depth: 3
    template: article-template.latex
    md_extensions: +raw_attribute
---

>>>>

---
title:  |
        | Legislature Size and Welfare: Evidence from Brazil
author: 
  - Umberto Mignozzetti[^umberto]
  - Gabriel Cepaluni[^gabriel]
  - Danilo Freire[^danilo]
date: May 28, 2022
thanks: We would like to thank Victor Araujo, Eric Arias, Mark Bennister, Taylor Boas, Natalia Bueno, Nick Cowen, Andrea Freitas, Saad Gulzar, Shoaib Jillani, Fernando Limongi, James Mahon, Lucas Mingardi, Lucas Novaes, Ingrid Oliveira, Nelson Ruiz, David Skarbek, Mike Slaven, Denis Stukal, the AJPS Editorial Team, and three anonymous reviewers for their valuable comments. We also thank the participants of seminars and conferences at the CEPESP-FGV, CUNY, INSPER, IPSA, IR-FGV, LACEA, LASA, Lincoln, MPSA, Notre Dame, NYU, PUC Chile, SPSA, and Uppsala. We gratefully acknolwedge funding from the São Paulo Research Foundation (FAPESP 2017/07290-5). Replication materials are available at <https://github.com/umbertomig/legislature-size-welfare>.
abstract: "How does legislature size impact public service provision? Despite the importance of institutional design for democratic governance, the effect of legislative features on citizen welfare remains little understood. In this paper, we use a formal model to show that increasing legislature size improves public goods delivery. We argue that changes in bargaining costs depend on whether additional legislators share the executive's party affiliation: More opposition members reduce the equilibrium public goods provision, while more government-aligned members increase it. We test this theory by exploiting sharp discontinuities in city council size in Brazil. We show that an additional city councilor has a 91% chance of belonging to the mayoral coalition, and this significantly improves primary school enrollment and infant mortality rates. To explore possible mechanisms, we surveyed 174 former city councilors and analyzed 346,553 bills proposed between 2005 and 2008. This paper has implications for the design of representative institutions."
abstractspacing: double
keywords: legislature size; legislative bargaining; local public goods provision; regression discontinuity design; Brazil
jelcodes: H41; H80; I31; I38
fontsize: 12pt
margin: 2cm
urlcolor: darkblue
linkcolor: Mahogany
citecolor: Mahogany
spacing: double
papersize: a4paper
bibliography: references.bib
biblio-style: apalike
output:
  pdf_document:
    citation_package: natbib
    fig_caption: yes
    number_sections: yes
    keep_tex: no
    toc: no
    toc_depth: 3
    template: article-template.latex
    md_extensions: +raw_attribute
---
